Sound Similarity, Sound Matching, and Differentiable Audio Synthesizer Control

### 2025-05-01 main.tex
abstract update

\begin{abstract}
Manual sound design with a synthesizer is an iterative process. When an artist has a target sound in mind, they listen to the output of their chosen synthesizer, determine whether the output is closer or farther from the target, and adjust the parameters until satisfied. There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict. Iterative sound-matching is the automation of this process, more formally, the continual programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity. We propose differentiable iterative sound-matching as the natural extension of the current available literature. 
Past works in sound-matching often relied on a narrow range of synthesizers in their analysis of ``best'' loss functions and overlooked the possible interactions between the loss and synthesizer parameters. Additionally, the results have rarely been verified with manual hearing tests. 
We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
For each loss function and synthesizer pair, we ran 300 experiments with randomized initial values and targets, and used the final parameter configurations, spectrogram distances, and manually assigned scores as key performance evaluation metrics of the performance. A post-hoc analysis of these performance measures showed that the best‐performing loss function varies based on the method of synthesis at hand, which highlights the value of novel similarity measures and de-prioritizes the search for the elusive universally optimal sound similarity measure.
\end{abstract}

>>>>

\begin{abstract}
Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied. There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict. Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function that quantifies similarity to the target sound. Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests. This leaves open whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preferences. We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning. To analyze the variability of loss function performance across synthesizers, we implemented differentiable subtractive, additive, and AM synthesizers, paired with four novel or established differentiable loss functions. For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials. We calculated parameter differences, spectrogram-distance metrics, and gathered blind listening scores. A post-hoc analysis shows that no single loss function dominates across all synthesis methods; instead, the optimal measure depends on the architecture at hand. These findings underscore the value of developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing a one-size-fits-all solution.
\end{abstract}
